Effects of bulk gas rotation on the emergent Lyman-$\alpha$ line in distant galaxies

### 2013-02-22 paper/Analisis.tex
Version_0.2 More text, and Fixed Pictures

\documentclass[usenatbib]{mn2e} 
\usepackage{amsmath} 
\usepackage{amssymb} 
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{epsfig}  
\def\be{\begin{equation}}
\def\ee{\end{equation}}
\def\ba{\begin{eqnarray}}
\def\ea{\end{eqnarray}}

@@

\title[Rotation in the Lyman-$\alpha$ line]{Effects of bulk gas rotation on
  the emergent Lyman-$\alpha$ line in distant galaxies}
\author[N. Garavito and J.E. Forero-Romero]{
\parbox[t]{\textwidth}{\raggedright 
  Nicolas Garavito.$^{1}$ 
  Jaime E. Forero-Romero$^{2}$ 
}
\vspace*{6pt}\\
$^{1}$Uni A
$^{2}$Uni B
}
\maketitle

@@

\section{Introduction}

@@

\section{Implementation of Bulk Gas Rotation}

We implement into CLARA the simplest model whereby a sphere rotates
with homogeneous angular velocity. We define a cartesian coordinate
system with its origin at the center of the sphere and the rotation
axis to be the $z$-axis, the components in the bulk velocity field, $\vec{v}
= v_{x}\hat{i} + v_{y}\hat{j} + v_{z}\hat{z}$. in the gas can be written as

@@

Additionaly in the postprocessing process we will take into account
different observer positions with respecto the cloud rotation
axis.

\begin{table}
\begin{center}
\begin{tabular}{ccc}\hline
Velocity (km s$^{-1}$) & $V_{\rm max}$&$0,\ 50,\ 100,\ 200,\ 300$\\
Hydrogen Optical Depth & $\tau_{H} $ & $10^{5},\ 10^{6},\ 10^{7}$\\
Dust Optical Depth & $\tau_{A}$ & XXX \\
Photons Distributions & & Central, Homogeneous\\
\hline
\end{tabular}
\caption{
Values for the varying input parameters in CLARA. Taking into account
all the possible combinations for these models
%odels, If we take into account all these possibles
% combinations we get 24 models with dust and 24 without dust for a
% total of 48 models 
} 
\label{table:models}
\end{center}
\end{table}

@@

Now if we take into account the position of the observer and compute
this for differents angles of observation $\theta$ the outcoming
spectra is modified as is shown in figure[2]. The main feature here is
that as the angle increases the line high decreases. It means that a
observer in the poles will observe a higher double peak than a
observer in the ecuator. We will understand this in more details when
we present the resaults of the scape fraction in section[].

\begin{figure}
\begin{center}
\includegraphics[scale=0.45]{Observers.png}
\end{center}
\caption{Spectra for different observers. Model:$V=300km/s$, Optical
  Depth $\tau=10^{7}$ and Central Distribution without dust.} 
\end{figure}

@@

In the previous section we describe the models that we study the
effect of rotation in the Lyman alpha line and see the effect of
rotation in the specrtums. Now we show the main resaults that we
obtain, in particual we pay spetial attention to the position of the
maximums, the width of the line and the escape fraction in function of
the rotational velocity and the angle of observation $\theta$.

In Figure[1] the effect that rotation have in the spectra is shown,
for both distributions central (Left) and homogeneous (right). This is
a general case in the sense that is independent of the position of an
external observer.

\subsection{Maximums Position}

\begin{figure}
    \includegraphics[width=0.45\textwidth]{0maximum7t.png}
    \includegraphics[width=0.45\textwidth]{0maximum6t.png}
    \includegraphics[width=0.45\textwidth]{0maximum5t.png}
  \label{figure:maxima}\caption{Width of the lyman-alpha line for all the models. }
\end{figure}
(comparar con la grafica del apendice de clara con el eje x el optical
depth y en el eje-y los maximos)

The maximus position gives information about the wave lenght of the
mayority of the outcoming photons after they interact with the gas, in
addition as a photon have more scatterings its wave lenght would be
larger than the initial which is $1216$ {\AA}. 
As we can see in Fig.[3] the position of the maximums does not change
with rotational velocity. But it change from homogeneous and central.

\subsection{Line Width}

%el ancho sirve para saber el rango de scattering que tienen los fotones.
Another effect that the rotation of the gas produce is in the  width
of the lyman alpha line. The width of the line provide information of
the the how many phtons scape with a particular wave lenght, in the
ideal case in which all the photons scape with the same wave lenght
the outcoming spectrum would be narrow. For all the models we study we
found that as the rotational velocity increase as the line width
increase this is show in Fig.[4].\\

\begin{figure*}
    \includegraphics[width=0.45\textwidth]{WidthCentral.png}
    \includegraphics[width=0.45\textwidth]{WidthHomogeneous.png}
    \includegraphics[width=0.45\textwidth]{Width6.png}
    \includegraphics[width=0.45\textwidth]{Width6HOM.png}
    \includegraphics[width=0.45\textwidth]{Width5.png}
    \includegraphics[width=0.45\textwidth]{Width5HOM.png}
  \label{figure:width}\caption{Width of the lyman-alpha line for all the models. }
\end{figure*}

@@

\subsection{Escape Fraction}

The fraction of photons that escape from de cloud of gas and dust is
defined as:\\

@@

Where NI is the initial number of photons and NF is the final. This
escape fraction is computed for all the models which resaults are
shown in Fig.[6]\\

@@

When the distribution is homogeneous the effect of velocity in the
escape fraction is clear while in the central model the effect is not
noterous. 
The main effect is that the escape fraction increase as the velocity
increase.

>>>>

\documentclass[usenatbib]{mn2e} 
\usepackage{amsmath} 
\usepackage{amssymb} 
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{epsfig} 
\usepackage{float} 
\def\be{\begin{equation}}
\def\ee{\end{equation}}
\def\ba{\begin{eqnarray}}
\def\ea{\end{eqnarray}}

@@

\title[Rotation in the Lyman-$\alpha$ line]{Effects of bulk gas rotation on
  the emergent Lyman-$\alpha$ line in distant galaxies}
\author[N. Garavito and J.E. Forero-Romero]{
\parbox[t]{\textwidth}{\raggedright 
  Nicolas Garavito-Camargo.$^{1}$ 
  Jaime E. Forero-Romero$^{2}$ 
}
\vspace*{6pt}\\
$^{1}$Uni A
$^{2}$Uni B
}
\maketitle

@@

\section{Introduction}
\label{sec:intro}

@@

\section{Implementation of Bulk Gas Rotation}
\label{sec:implementation}

We implement into CLARA the simplest model whereby a sphere rotates
with homogeneous angular velocity. We define a cartesian coordinate
system with its origin at the center of the sphere and the rotation
axis to be the $z$-axis, the components in the bulk velocity field, $\vec{v}
= v_{x}\hat{i} + v_{y}\hat{j} + v_{z}\hat{k}$. in the gas can be written as

@@

Additionaly in the postprocessing process we will take into account
different observer positions with respect to the cloud rotation
axis.

\begin{table}
\begin{center}
\begin{tabular}{ccc}\hline
Velocity (km s$^{-1}$) & $V_{\rm max}$&$0,\ 50,\ 100,\ 200,\ 300$\\
Hydrogen Optical Depth & $\tau_{H} $ & $10^{5},\ 10^{6},\ 10^{7}$\\
Dust Optical Depth & $\tau_{A}$ & XXX \\
Photons Distributions & & Central, Homogeneous\\
\hline
\end{tabular}
\caption{
Values for the varying input parameters in CLARA. Taking into account
all the possible combinations for these models
} 
\label{table:models}
\end{center}
\end{table}

@@

Now if we take into account the position of the observer and compute
this for differents angles of observation $\theta$ the outcoming
spectra is modified as is shown in Figure~\ref{DifferentObservers}. The main feature here is that as the angle increases the line height decreases. It means that a observer in the poles will observe a higher double peak than a
observer in the ecuator. We will understand this in more detail when
we present the results of the scape fraction in section~\ref{sec:EF}.

\begin{figure}
\begin{center}
\includegraphics[scale=0.45]{Observers.png}
\end{center}
\caption{Spectra for different observers. Model:$V=300km/s$, Optical
  Depth $\tau=10^{7}$ and Central Distribution without dust.} 
  \label{DifferentObservers}
\end{figure}

@@

In the previous section we describe the models that we use to study the
effect of rotation in the Lyman alpha line for this we saw the outcoming specrtums including rotation. Now we show the main results that we
obtain, in particual we pay spetial attention to the position of the
maximums, the width of the line and the escape fraction in function of
the rotational velocity and the angle of observation $\theta$.

In Figure~\ref{fig:HOMandCentral} the effect that rotation have in the spectra is shown,
for both distributions central (Left) and homogeneous (right). This is
a general case in the sense that is independent of the position of an
external observer.

\subsection{Maxima Positions}
\label{sec:MP}

\begin{figure}
    \includegraphics[width=0.45\textwidth]{maximumvsODDifSpeedsCentral.png}
    \includegraphics[width=0.45\textwidth]{maximumvsODDifSpeedsHOM.png}
  \label{figure:maximavsOD}\caption{Position of the maxima in the outgoing spectra for differents Optical Depths, (up)Central Distriution, (Down) Homogeneous Distribution.}
\end{figure}

\begin{figure}
    \includegraphics[width=0.45\textwidth]{maximumvsvelocitiesDifODCentral.png}
    \includegraphics[width=0.45\textwidth]{maximumvsvelocitiesDifODHOM.png}
  \label{figure:maximumsvsvelocities}\caption{Position of the maxima in the outgoing spectra for differents Rotational velocities, (up)Central Distriution, (Down) Homogeneous Distribution.}
\end{figure}

The maximums position gives information about the wave lenght of the
mayority of the outcoming photons after they interact with the gas, in
addition as a photon have more scatterings its wave lenght would be
larger than the initial which is $1216$ {\AA}. 
As we can see in Figure ~\ref{figure:maximumsvsvelocities} the position of the maximums does not change with rotational velocity for the central dsitribution. On the other hand for the homogeneous distriution the maxima position $X_{m}$ change for $Log_{10}\tau=5$, this is because at higher rotational velocities photons escape with less scatterings, for $v=100\dfrac{km}{s}$ the majority of the photons still escape with scatterings but for $v=200\dfrac{km}{s}$ the number of photons that escape with a few or any scatterings is higher, this is why the maxima change to the center of the spectra.\\

If we now study the effect of the optical depth $\tau$ in the maxima position $X_{m}$ Figure~\ref{figure:maximavsOD}, we found that as the optical depth increase (include reference of previous articles) the maxima position increase and this is a well known result, but when rotation is included the dependence is not linear.
%(think a lit bit more on this, maybe make a regression)
%explain the method to find the maxima position

\subsection{Line Width}
\label{sec:LW}

Another effect that the rotation of the gas produce is in the  width
of the lyman alpha line. The width of the line provide information of
how many photons scape with a particular wave lenght, in the
ideal case in which all the photons scape with the same wave lenght
the outcoming spectrum would be narrower. For all the models we study we
found that as the rotational velocity increase the line width
increase this is show in Fig.[4].\\

\begin{figure*}
    \includegraphics[width=0.45\textwidth]{Width7Central.png}
    \includegraphics[width=0.45\textwidth]{Width7Homogeneous.png}
    \includegraphics[width=0.45\textwidth]{Width6.png}
    \includegraphics[width=0.45\textwidth]{Width6HOM.png}
    \includegraphics[width=0.45\textwidth]{Width5.png}
    \includegraphics[width=0.45\textwidth]{Width5HOM.png}
  \label{figure:width}\caption{Width of the lyman-alpha line for all the models. }
\end{figure*}

@@

\subsection{Escape Fraction}
\label{sec:EF}

The fraction of photons that escape from the cloud of gas and dust is
defined as:\\

@@

Where NI is the initial number of photons and NF is the final. This
escape fraction is computed for all the models which results are
shown in Fig.[6]\\

@@

When the distribution is homogeneous the effect of velocity in the
escape fraction is clear while in the central model the effect is not
notorious. 
The main effect is that the escape fraction increase as the velocity
increase.
